### 2016-03-05 document/table_meanfpp.tex
table

\begin{deluxetable}{cccc}
\tabletypesize{\scriptsize}
\tablecaption{Mean FPPs of candidate KOIs with reliable \vespa\ calculations
\tablabel{meanFPP} }
\tablehead{\colhead{Selection} & \colhead{Number} & \colhead{Mean FPP}}
\startdata
all & 2775 & 0.131 \\
singles & 1623 & 0.170 \\
multis & 1152 & 0.076 \\
$R_p > 15\,R_\oplus$ & 208 & 0.746 \\
$10\,R_\oplus < R_p < 15\,R_\oplus$ & 75 & 0.245 \\
$4\,R_\oplus < R_p < 10\,R_\oplus$ & 222 & 0.221 \\
$2\,R_\oplus < R_p < 4\,R_\oplus$ & 935 & 0.075 \\
$R_p < 2\,R_\oplus$ & 1335 & 0.053 
\enddata
\end{deluxetable}