### 2016-03-07 document/table_meanfpp.tex
tables updated

\begin{deluxetable}{cccc}
\tabletypesize{\scriptsize}
\tablecaption{Mean FPPs of candidate KOIs with reliable \vespa\ calculations
\tablabel{meanFPP} }
\tablehead{\colhead{Selection} & \colhead{Number} & \colhead{Mean FPP} 
\startdata
all & 2791 & 0.133 \\
singles & 1635 & 0.177 \\
multis & 1156 & 0.071 \\
$R_p > 15\,R_\oplus$ & 240 & 0.820 \\
$10\,R_\oplus < R_p < 15\,R_\oplus$ & 90 & 0.212 \\
$4\,R_\oplus < R_p < 10\,R_\oplus$ & 241 & 0.185 \\
$2\,R_\oplus < R_p < 4\,R_\oplus$ & 1144 & 0.050 \\
$R_p < 2\,R_\oplus$ & 1076 & 0.050 
\enddata
\end{deluxetable}