### 2016-03-15 document/table_meanfpp.tex
nb and figure work

\begin{deluxetable}{cccc}
\tabletypesize{\scriptsize}
\tablecaption{Mean FPPs of candidate KOIs with reliable \vespa\ calculations
\tablabel{meanFPP} }
\tablehead{\colhead{Selection} & \colhead{Number} & \colhead{Mean FPP} } 
\startdata
all & 2857 & 0.154 \\
singles & 1688 & 0.206 \\
multis & 1169 & 0.080 \\
$R_p > 15\,R_\oplus$ & 256 & 0.837 \\
$10\,R_\oplus < R_p < 15\,R_\oplus$ & 91 & 0.220 \\
$4\,R_\oplus < R_p < 10\,R_\oplus$ & 252 & 0.218 \\
$2\,R_\oplus < R_p < 4\,R_\oplus$ & 1160 & 0.065 \\
$R_p < 2\,R_\oplus$ & 1098 & 0.069 
\enddata
\end{deluxetable}